Solving \\ the \\ 15-Puzzle \\ \& \\ Rubik's Cube

### 2022-09-11 latex/Chapters/SlidingPuzzle.tex
update

\Subsection{Solvers' comparison}
\label{ssec:33SPSC}
\label{sec:S33DRLDQL}
Let me discuss a comparison of several algorithms on 1000 random puzzles generated for a number of random shuffling (with best-effort-no-backtracking) from 0 to 50 in step of 2, as well as for perfect shuffling (denoted by $\infty$) on the comparison graphs. The results are shown in figure \ref{fig:33SPPerformance}
\\
DL: 100 seq 15 to 31 shuffles 10k epochs connected 600 300 100
\\
Manhattan vs ++
\\
DQL vs DRL
\\
DxL vs Perfect
\\ Put table for nb\_shuffle=inf and highlight best in diff categories (cost, optimality score, run time, nodes)
\\ say how much of the puzzles the solvers have seen

\begin{figure}[H]
  \noindent
  \makebox[\textwidth]{\includegraphics[scale=0.6]{./Figures/33SPPerformance}}
  \caption[33SPPerformance]{Solvers' performance comparison 3x3 \textbf{SP}}
  \label{fig:33SPPerformance}
\end{figure}

>>>>

\Subsection{Solvers' comparison}
\label{ssec:33SPSC}
\label{sec:S33DRLDQL}
As promised earlier in table \ref{tab:gridSP}, I have run the Solver's comparison test with 10 different solvers (Naive, \textbf{MCTS}, \textbf{BFS}, A$^{*}$ with the following heuristics: Manhattan, Manhattan++, Perfect, \textbf{DRL}, \text{DQL} and \text{DL}), to get an idea of their respective success rate, strengths and weaknesses. The \textbf{DRL} and \textbf{DQL} heuristics were trained with the exact same parameters and a fully connected network of 3 hidden layers with 600, 300 and 100 nodes. For the \textbf{DL} heuristic (trained on perfect solutions from A$^{*}$[Perfect]) I tried with the same fully connected network as well as with a combination of parallel fully connected and convolutional layers (see figure \ref{fig:33SPnets} for the exact architecture I used). The detailed results (percentage of configurations solved, optimality score, median and max cost, run time and expanded nodes) can be found on the next page  in figure \ref{fig:33SPPerformance}. Each solver had to solve 1,000 random puzzles for each level of difficulty (scrambling from goal) going from 5 to 50 in step of 5, as well as for perfect shuffling. Let me list some of the important findings from this rather comprehensive experiment:

\begin{itemize}
\item As expected, \textbf{BFS} becomes quickly hopeless.
\item The naive solver is obvioulsy by far the fastest but finds poor solutions (about 2.5 longer on average than optimal).
\item As discussed in the previous section, \textbf{MCTS} is slow, and even with much exploration and \textbf{BFS} tree trimming found solution about 25\% longer than optimal.
\item I originally was naively hoping a pure convolutional network would do well but did not manage to get the loss to converge. I later realized that this is quite different from the typical image patterns recognition as there is less direct translation and scale invariants at play here, and decided instead to try combining convolutional with fully connected layers in parallel. That improved things, but overall I saw little difference in performance from my simpler 3-hidden-layer fully connected network.
\item \textbf{DRL} and \textbf{DQL} heuristics, though not as efficient in terms node-expansion as the perfect heuristic and the \textbf{DL} heuristics, were rather surprisingly doing better than even Manhattan++. They ended up \textit{seeing} around 25\% of all the possible configurations during training (in an \textit{unsupervised} way though) so did show some reasonable amount of generalization out of sample. Their optimality score was an impressive 99\%+ at all levels of difficuly for \textbf{DRL} and 98\%+ at all levels of difficuly for \textbf{DQL}. Even more interestingly, the \textbf{DQL} heuristic expanded almost 4 times less nodes than \textbf{DRL}, suggesting some benefit from the joint learning of the policy and value function.
\item The \textbf{DL} heuristics (fully connected and convolutional) performed almost on par with the perfect heuristic (optimality score of 99\%+ at all levels of difficulty and nodes expansion only increased by about 12\%), despite seeing only 8\% of the configurations by the end of their training.
\end{itemize}

\begin{figure}
  \noindent
  \makebox[\textwidth]{\includegraphics[scale=0.45]{./Figures/33SPnets}}
  \caption[33SPnet]{Architecture used for the \textbf{DL} heuristics training on the 3x3 \textbf{SP}}
  \label{fig:33SPnets}
\end{figure}

\begin{figure}[H]
  \noindent
  \makebox[\textwidth]{\includegraphics[scale=0.55]{./Figures/33SPPerformance}}
  \caption[33SPPerformance]{Solvers' performance comparison 3x3 \textbf{SP}}
  \label{fig:33SPPerformance}
\end{figure}
